---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rELAN

<!-- badges: start -->

<!-- badges: end -->

## Overview

rELAN provides a tool to import ELAN files (.eaf), which are generated by the annotation software ELAN, directly into R as `data.frame`s.

## Installation

You can install rELAN with:

```{r installation, eval=FALSE}
# install.packages("devtools")
devtools::install_github("relan-package/rELAN")
```

## Usage

Depending on your working directory, the first argument is the file name, or path and file name, as a string. `wide_format = TRUE` returns a wide `data.frame` where each tier has its own column and rows are merged, so `ANNOTATION_VALUE`s replace `NA`s. This will also have fewer data than the original extracted `data.frame`.

```{r examples, eval=FALSE}
library(rELAN)

exctract_annotations("ELAN_files/frog_story.eaf")

extract_annotations("ELAN_files/pear_story.eaf",
                    wide_format = TRUE)

```

## Why rELAN / more Information

So far, one of the most common ways to import the annotation data into R, was by a two step process. First, you needed to use ELAN's function to export the ELAN file, which is written in XML, as a tab-delimited text, for instance. This exported file could then be imported into R as a `data.frame`. Thus, using rELAN has three advantages:

1.  The import is a single step, which is more economical in general.
2.  If you need to add, change, or delete annotations, you only need to modify the ELAN file and import it into R again instead of changing the ELAN file, the tab-delimited file, and importing it into R.
3.  Importing with rELAN delivers you every information of the ELAN file concerning the annotations. The tab-delimited text file only contains limited data relating to the annotations.

The default of `extract_annotations()` produces a long `data.frame` with all data relating to the annotations. However, you can get a `data.frame` with fewer data, where each tier has its own column, by using the argument `wide_format = TRUE`.

While there are other packages available in R and Python for working with ELAN files directly, rELAN stands out for its ability to calculate the time values from reference annotations.
